---
title: "transcriptomics_plots"
author: "Degnan, David J"
date: "2023-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
library(data.table)
library(plotly)
library(DT)
```

## Read data 

```{r}
transcriptomics <- fread("~/Downloads/transcriptomics_out.txt")
```

## Chemical Page Level

#### Counts of Significant Genes per Concentration

```{r}
# As an example, pull chemical
ChemicalID <- "50-32-8"

# Data at the chemical level
chem_level <- transcriptomics[transcriptomics$CASNO == ChemicalID,]

chem_level %>%
  filter(adj_p_value <= 0.05) %>%
  count(Concentration) %>%
  mutate(Concentration = as.factor(Concentration)) %>%
  plot_ly(x = ~Concentration, 
          color = ~Concentration,
          y = ~n,
          type = "bar") %>%
  layout(yaxis = list(title = "Number of Significant Genes"),
         xaxis = list(title = "Concentration in uM"),
         legend = list(title = list(text = "Concentration")))
```

Clicking on this bar chart will trigger changes for the significantly expressed genes.
In other words, all of the following plots in this section "Chemical Page Level"
should change depending on the selected concentration, with the exception of table of 
all genes. 

#### Expression of significant genes 

```{r}
# Concentration selection
Concentration <- 1

chem_level %>%
  filter(Concentration == 1) %>%
  mutate(
    flag = factor(ifelse(adj_p_value <= 0.05, 1, 0) * ifelse(Log2FoldChange >= 0, 1, -1), levels = c(-1, 0, 1)), 
    adj_p_value = -1 * log10(adj_p_value)
  ) %>%
  plot_ly(x = ~Log2FoldChange,
          y = ~adj_p_value,
          color = ~flag) %>%
  layout(yaxis = list(title = "-Log10 P-Value"),
         xaxis = list(title = "Log2 Fold Change"),
         legend = list(title = list(text = "Direction")),
         title = list(text = "Selected Concentration is 1uM"))
```

#### Table of all genes 

```{r}
chem_level %>%
  filter(Concentration == 1) %>%
  dplyr::select(-c(condition)) %>%
  dplyr::rename(`Gene ID` = GeneID, `Log2 Fold Change` = Log2FoldChange, `P Value` = adj_p_value, Flag = flag) %>%
  datatable(options = list(scrollX = TRUE))
```

Clicking a row on the table will open a "Gene Page". 


